---
title: "Breast Cancer Prediction"
author: "Mansour Alhumaid"
date: "28 Nov 2020"
output: html_document
---


```{r, include = FALSE}
knitr::opts_chunk$set(cache = TRUE, 
                      message = FALSE,
                      warning = FALSE,
                      echo = FALSE,
                      autodep = TRUE, 
                      fig.align = "center")
library(tidyverse)
library(rio)
library(ggstatsplot)
library(rsample)   # data splitting
library(ggplot2)   # plotting
library(earth)     # fit MARS models
library(caret)     # automating the tuning process
library(vip)       # variable importance
library(pdp)       # variable relationships
library(randomForest) #its works in regression and classification
library( kableExtra)
library(corrplot)
library(caTools)
library(car)
#library(FSelector)
library(data.tree)
library(rpart)
#library(raprt.plot)
library(class)



```


In this study, a dataset contain Clinical features were observed or measured for 64 patients with breast cancer and 52 healthy controls. The dataset is downloaded from The UCI Machine Learning Repository
There are 10 predictors, all quantitative, and a binary dependent variable, indicating the presence or absence of breast cancer. The predictors are anthropometric data and parameters which can be gathered in routine blood analysis. Prediction models based on these predictors, if accurate, can potentially be used as a biomarker of breast cancer.

Data From:
https://archive.ics.uci.edu/ml/datasets/Breast+Cancer+Coimbra



```{r}

# read the data 
df <- read.csv("Data/dataR2.csv")
```




#### EDA
as start on the exploration process on the data, to explore available features and observations of each:


```{r}
nrow(df)
ncol(df)
head(df)
tail(df)
str(df)

```

#### Conversion
```{r}

df$Classification=factor(df$Classification)
df$Age=factor(df$Age)
df=subset(df,select = -c(1,10))
```





#### finding null values
```{r}
which(is.na(df))
any(is.na(df))
```




####  Visualization
```{r}
boxplot(df$MCP.1)
```




